# Overall Disease Burden

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data/burden-of-disease-all-causes.csv"))
head(data_overall)

```

```{r}
# Filter the dataset to include only four countries.

data_overall_mini <- data_overall %>% 
  filter(Entity %in% c("Ghana", "China","Ethiopia","India"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries. 

# You should pivot the data wider to show each country in a separate column.
# Ensure tidyr is loaded
library(tidyr)

# Pivot the data wider

data_overall_wide <- data_overall_mini %>%
  pivot_wider(
    names_from = Entity, 
    values_from = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`
  )

#Remove code column
data_overall_wide1 <- subset(data_overall_wide, select = -Code)

# Rename columns
data_overall_wide1 <- rename(data_overall_wide1, 
                             `DALY/China` = China, 
                             `DALY/Ethiopia` = Ethiopia, 
                             `DALY/Ghana` = Ghana, 
                             `DALY/India` = India)


kable(data_overall_wide1)



  
# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 

library(ggplot2)

plote_data_mini <- ggplot(data_overall_mini, aes(x=Year, y=`DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`, colour = Entity, group=Entity))+
  geom_line(aes(linewidth = 1))+
  labs(
    title = "Disability-Adjusted Life Years (DALYs) by Entity",
    x = "Year",
    y = "DALYs (Age-standardized Rate)"
  ) +
  theme_minimal()

plot(plote_data_mini)


```
```

## Summary of Overall Disease Burden Findings

China has the lowest DALY overtime and Ethiopia recorded the highest DALY overtime. 
It is also observed that all the four country's DALY declined steadily overtime with Ethiopia having the steepest decline.
And the highest DALY was recorded in 1990 by Ethiopia
